Analysis of the Anomalous Doppler Effect from Quantum Theory to Classical Dynamics Simulations

### 2025-12-02 paper modification/CPB-tex-template/CPB manuscript/Latex_manuscript/paper_v2.tex
new materials
revising

%=================================================================================================%
% This is a template designed for Chin. Phys. B (Dated 12 November 2014)
% Only one step to compile: PDFTeXify
%=================================================================================================%
\documentclass{cpbtex3}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{authblk}
%\usepackage[backend=biber]{biblatex}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{titlesec}      % For customizing section titles
% ===== 1. Small font for affiliations =====
\renewcommand{\Affilfont}{\small} % Set affiliation font size to small

@@

\section{Introduction}
The Anomalous Doppler Effect (ADE) \cite{tamm1959general,frank1960optics,ginzburg1960certain,shustin1971transformation}, in which the observed frequency shift behaves contrary to the conventional Doppler Effect under specific conditions, was first theoretically predicted by Soviet physicist Vitaly L. Ginzburg \cite{ginzburg1946radiation}. This phenomenon arises when a system moves with a velocity exceeding the phase velocity of light in the medium, transferring its translational kinetic energy into internal energy while emitting radiation. A notable example, discussed by Frank in his 1958 Nobel lecture \cite{frank1960optics}, shows that radiation emission does not occur through the typical transition from an excited state to a lower energy state. Instead, it proceeds from a lower to a higher energy state, powered by the system's translational kinetic energy. This counterintuitive prediction has attracted considerable attention and has inspired extensive research \cite{nezlin1976negative,santini1984anomalous,kho1988slow,wang2016multi,guo2018control,liu2018role,shi2018superlight,filatov2021role}.

@@

\section{Conclusion}\label{sec:Conclusion}
This paper presents a simple yet effective method for analyzing the resonant processes associated with the Normal Doppler Effect (NDE) and Anomalous Doppler Effect (ADE). By combining quantum theory with angular momentum conservation analysis, it is shown that the parameter \( m \) in the resonance condition  
\(
\omega = \vec{k} \cdot \vec{v} + m\omega_{ce}
\)
is directly associated with the angular momentum of the resonant wave. Numerical simulations based on the Volume Preserving Algorithm (VPA) further support the quantum results, confirming both the angular momentum correspondence of \( m \) and the energy transfer characteristics. In future work, the interaction between electron energy transformation and helicon waves will be investigated within a plasma environment, aiming to provide deeper insights into applications such as plasma heating and runaway electron suppression.
\section{Acknowledgements}
This work is supported by National Magnetic Confinement Fusion Energy Program of China under Contract No. 2019YFE03020001
%\addcontentsline{toc}{chapter}{Appendix A: Appendix section heading}
\section*{Appendix A:Classical Analysis of Anomalous Doppler Resonance}
Neglecting the static electric field and relativistic effect , we provide a brief derivation of the energy transformation process based on classical dynamical equations:
\begin{align}
m_e \frac{d\vec{v}_\parallel}{dt} &= -e(\vec{v}_\perp \times \vec{B}_\perp) \label{eq:19} \\
m_e \frac{d\vec{v}_\perp}{dt} &= -e(\vec{v}_\perp \times \vec{B}_0 + \vec{v}_\parallel \times \vec{B}_\perp + \vec{v}_\perp \times \vec{B}_0) \label{eq:20}
\end{align}

>>>>

%=================================================================================================%
% This is a template designed for Chin. Phys. B (Dated 12 November 2014)
% ompile: PDFTeXify
%=================================================================================================%
\documentclass{cpbtex3}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{authblk}
%\usepackage[backend=biber]{biblatex}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{titlesec}      % For customizing section titles
% ===== 1. Small font for affiliations =====
\renewcommand{\Affilfont}{\small} % Set affiliation font size to small

@@

\section{Introduction}
The Anomalous Doppler Effect (ADE) \cite{tamm1959general,frank1960optics,ginzburg1960certain,shustin1971transformation}, in which the observed frequency shift behaves contrary to the conventional Doppler Effect under specific conditions, was first theoretically predicted by Soviet physicist Vitaly L. Ginzburg \cite{ginzburg1960certain}. This phenomenon arises when a system moves with a velocity exceeding the phase velocity of light in the medium, transferring its translational kinetic energy into internal energy while emitting radiation. A notable example, discussed by Frank in his 1958 Nobel lecture \cite{frank1960optics}, shows that radiation emission does not occur through the typical transition from an excited state to a lower energy state. Instead, it proceeds from a lower to a higher energy state, powered by the system's translational kinetic energy. This counterintuitive prediction has attracted considerable attention and has inspired extensive research \cite{nezlin1976negative,santini1984anomalous,kho1988slow,wang2016multi,guo2018control,liu2018role,shi2018superlight,filatov2021role}.

@@

\section{Conclusion}\label{sec:Conclusion}
This paper presents a simple yet effective method for analyzing the resonant processes associated with the Normal Doppler Effect (NDE) and Anomalous Doppler Effect (ADE). By combining quantum theory with angular momentum conservation analysis, it is shown that the parameter \( m \) in the resonance condition  
\(
\omega = \vec{k} \cdot \vec{v} + m\omega_{ce}
\)
is directly associated with the angular momentum of the resonant wave. Numerical simulations based on the Volume Preserving Algorithm (VPA) further support the quantum results, confirming both the angular momentum correspondence of \( m \) and the energy transfer characteristics. In future work, the interaction between electron energy transformation and helicon waves will be investigated within a plasma environment, aiming to provide deeper insights into applications such as plasma heating\cite{zhao2022plasma}  and runaway electron suppression.
\section{Acknowledgements}
This work is supported by National Magnetic Confinement Fusion Energy Program of China under Contract No. 2019YFE03020001
%\addcontentsline{toc}{chapter}{Appendix A: Appendix section heading}
\section*{Appendix A:Classical Analysis of Anomalous Doppler Resonance}
Neglecting the static electric field and relativistic effect , we provide a brief derivation of the energy transformation process based on classical dynamical equations:
\begin{align}
m_e \frac{d\vec{v}_\parallel}{dt} &= -e(\vec{v}_\perp \times \vec{B}_\perp) \label{eq:19} \\
m_e \frac{d\vec{v}_\perp}{dt} &= -e(\vec{v}_\perp \times \vec{B}_0 + \vec{v}_\parallel \times \vec{B}_\perp + \vec{v}_\perp \times \vec{B}_0) \label{eq:20}
\end{align}
